### 2018-07-23 manuscript.tex
merging internal review

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% LIVECOMS ARTICLE TEMPLATE FOR BEST PRACTICES GUIDE
%%% ADAPTED FROM ELIFE ARTICLE TEMPLATE (8/10/2017)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PREAMBLE
\documentclass[9pt,tutorial]{livecoms}
% Use the 'onehalfspacing' option for 1.5 line spacing
% Use the 'doublespacing' option for 2.0 line spacing
% Use the 'lineno' option for adding line numbers.
% The 'bestpractices' option for indicates that this is a best practices guide.
% Omit the bestpractices option to remove the marking as a LiveCoMS paper.
% Please note that these options may affect formatting.

\usepackage{lipsum} % Required to insert dummy text
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\DeclareSIUnit\Molar{M}
\usepackage[italic]{mathastext}
\graphicspath{{figures/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% IMPORTANT USER CONFIGURATION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\versionnumber}{0.1} % TODO: can we extract this number from a git tag?
\newcommand{\githubrepository}{\url{github.com/markovmodel/pyemma_tutorials}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Introduction to Markov state modeling with the PyEMMA software --- v\versionnumber}

\author[1\authfn{1}*]{Christoph Wehmeyer}
\author[1\authfn{1}]{Martin K. Scherer}
\author[1\authfn{1}]{Tim Hempel}
\author[1]{Simon Olsson}
\author[2]{Brooke E. Husic}
\author[1*]{Frank Noé}
\affil[1]{Department of Mathematics and Computer Science, Freie Universität Berlin, Arnimallee 6, 14195 Berlin, Germany}
\affil[2]{Department of Chemistry, Stanford University, Stanford, California 94305, USA}

\corr{christoph.wehmeyer@fu-berlin.de}{CW}
%\corr{m.scherer@fu-berlin.de}{MKS}
%\corr{tim.hempel@fu-berlin.de}{TH}
%\corr{simon.olsson@fu-berlin.de}{SO}
%\corr{brookehusic@gmail.com}{BEH}
\corr{frank.noe@fu-berlin.de}{FN}

\contrib[\authfn{1}]{These authors contributed equally to this work}

\blurb{This LiveCoMS document is maintained online on GitHub at \githubrepository{}; to provide feedback, suggestions, or help improve it, please visit the GitHub repository and participate via the issue tracker.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}
\maketitle

\begin{abstract}
This tutorial provides an introduction to the construction of Markov models of molecular kinetics from molecular dynamics (MD) trajectory data.
Using Jupyter notebooks, we will guide the user through the basic functionality as well as the more common advanced mechanisms.
Short exercises to self check the learning progress and a notebook on troubleshooting complete this basic introduction.
\end{abstract}

\end{frontmatter}

\section{Introduction}

PyEMMA~\cite{pyemma} (\url{http://emma-project.org}) is a software for the analysis of molecular dynamics (MD) simulations using Markov state models~\cite{schuette-msm,singhal-msm-naming} (MSMs).
The package is written in Python (\url{http://python.org}), relies heavily on NumPy/SciPy~\cite{numpy,scipy}, and is compatible with the scikit-learn~\cite{sklearn} framework for machine learning.

\subsection{Scope}

In this tutorial, we assume the reader's familiarity with the basic theory behind the MSM approach (see Sec.~\ref{sec:background}) and focus on usage of PyEMMA.
Nevertheless, we will mention important theoretical concepts when appropriate throughout the tutorial.
We also assume the reader is familiar with molecular dynamics analyses of proteins and peptides and commonly used structural components of these systems; for a review of the simulation of biomolecular macromolecules, see~\cite{dror2012biomolecular}.

The tutorial is divided into lessons on specific topics, each accompanied by a Jupyter~\cite{jupyter} notebook containing code, instructions, and exercises.
The lessons start with a showcase of the PyEMMA workflow and follow up with in-depth lessons on specific topics.

\section{Prerequisites}

In the following, we summarize the recommended theoretical background knowlegde of Markov state modeling for this tutorial.
Then, we address the software requirements to work through the lessons.

\subsection{Background knowledge}
\label{sec:background}

For those unfamiliar with Markov state modeling, ``\emph{Markov State Models: From an Art to a Science}''~\cite{msm-brooke} provides a recent overview, while ``\emph{Markov models of molecular kinetics: Generation and validation}''~\cite{msm-jhp} describes the basic MSM theory and methodology in detail..

In addition to publications on theory and application of Markov state modeling~\cite{msm-book,buchete-msm-2008,noe-tmat-sampling,bowman-msm-2009,noe-folding-pathways,sarich-msm-quality,noe-fingerprints,noe-dy-neut-scatt,Chodera2014,ben-rev-msm,simon-mech-mod-nmr,oom-feliks},
we also recommend the literature on time-lagged independent component analysis (TICA)~\cite{tica,tica3,tica2,kinetic-maps}, transition path theory (TPT)~\cite{weinan-tpt,metzner-msm-tpt},
hidden Markov state models (HMMs)~\cite{noe-proj-hid-msm,hmm-baum-welch-alg,hmm-tutorial}, and variational techniques~\cite{noe-vac,vamp-preprint,gmrq}, as these topics play important roles within the standard MSM workflow.

\subsection{Software/system requirements}

\begin{figure}
\includegraphics[width=0.48\textwidth]{figure_1}
\caption{The PyEMMA workflow: The MD trajectories are processed and discretized (first row). A Markov model is estimated from the resulting discrete trajectories and validated (middle row). By iterating between data processing and MSM estimation / validation, a valid model is obtained that can be analyzed (last row).}
\label{fig:workflowchart}
\end{figure}

We utilize Jupyter~\cite{jupyter} notebooks to show code examples along with figures and interactive widgets to display molecules.
The user can install all necessary packages in one step using the \texttt{conda} command provided by the Anaconda Python stack (\url{https://anaconda.com}).
We rely on Anaconda, because it resolves and installs dependencies, and provides pre-compiled versions of hard to build packages like SciPy.

The installation contains a launcher command to start the Jupyter notebook server as well as the notebook files.
The data for the demonstrated test systems is downloaded upon the first use and is cached for future invocations of the tutorial.

The underlying software stack for running the tutorial importantly consists of:
\begin{itemize}
\item NumPy -- vectorized operations on multi-dimensional arrays~\cite{numpy}
\item SciPy -- linear algebra, sparse computations~\cite{scipy}
\item Matplotlib -- Visualization~\cite{matplotlib}
\item \textbf{PyEMMA} -- MSM/HMM estimation, validation, analysis, and visualization~\cite{pyemma}
\item mdtraj -- reading MD simulation data, calculate observables from MD~\cite{mdtraj}
\end{itemize}

The software is currently supported for Python versions $3.5$ and $3.6$ on the operating systems Linux, OSX, and Windows.

Should the user prefer to not use Anaconda, a manual installation via the pip installer is possible.
Alternatively, one can use the Binder service to view and run the tutorials online in any browser.

\section{Content and links}

This tutorial consists of nine Jupyter notebooks which introduce the basic features of PyEMMA.
The first notebook (00), which we will summarize in the following, showcases the entire estimation, validation, and analysis workflow.
The goal of this introductory notebook (00) is to provide the user with the typical steps required to obtain a validated MSM analysis of protein or peptide simulation data.
The eight subsequent notebooks (01--08) provide in-depth lessons on specific topics, an introduction to HMMs, and guidelines how to deal with common problems during an MSM estimation.

\subsection{The PyEMMA workflow}

In short, the workflow consists of extracting molecular features from the raw data (01) which we transform into a suitable, low-dimensional subspace and discretize (02).
Then, we estimate a maximum likelihood or Bayesian MSM from the discrete trajectories and perform validation tests (03).
We can then analize the stationary and kinetic properties of the MSM (04),
find metastable macrostates and apply transition path theory (TPT) to identify the pathways of conformation change (05),
and compute expectation values for experimental observables (06).

In order to guide the user through the process before going into technical details,
the first notebook goes through the full analysis pipeline for,
we analyze the backbone conformation dynamics of a pentapeptide (Trp-Leu-Ala-Leu-Leu, Fig.~\ref{fig:io-to-ck}a) using 25 independently simulated trajectories with implicit solvent and a trajectory saving interval of $0.1$ ns~\cite{pyemma}.

\subsection{Estimation and validation}

\begin{figure}
\includegraphics{figure_2}
\caption{Exemplary analysis of the conformational dynamics of a pentapeptide backbone: (a) The Trp-Leu-Ala-Leu-Leu pentapeptide in licorice representation.
(b) The VAMP-2 score indicates which of the tested featurizations contains the highest kinetic variance.
(c) The  sample density projected of the onto the first two independent components of a TICA transformation at lag time $\tau=0.5$ ns shows multiple density maxima and
(d) the timeseries of the first two independent components show rare transition events.
(e) The convergence behavior of the first four implied timescales indicates that a lag time of $\tau=0.5$ ns is suitable for estimating an MSM;
the shaded areas indicate $95\%$ confidence intervals. (f) A Chapman-Kolmogorov test shows that an MSM estimated at lag time $\tau=0.5$ ns under the assumption of four metastable states accurately predicts the kinetic behavior on longer timescales;
the shaded areas indicate $95\%$ confidence intervals.}
\label{fig:io-to-ck}
\end{figure}

To begin, we start with the raw trajectory data, and we seek to constuct and validate an MSM, which is represented by a transition matrix.

It is initially unknown which features (e.g. distances, contact maps, angles, etc.) are good descriptors of the slow CVs, i.e., are best suited to resolve the long-time dynamics~\cite{NoeClementiReview}.
A systematic way of making this choice has been introduced through the variational principle of conformation dynamics~\cite{noe-vac,nueske-vamk} and recently the more general version of nonequilibrium data, VAMP~\cite{vamp-preprint}.
VAMP provides a score that can be calculated for each feature in order to find good indicators of the long-time dynamics.
This score is computed via cross-validation in order to balance underfitting (too few features) and overfitting (too many features)~\cite{gmrq}.
Here we utilize the VAMP-2 score, which maximizes the kinetic variance of the estimated dynamical operator~\cite{kinetic-maps}.
To choose among three different molecular features relevant to protein structure, we compute the (cross-validated) VAMP-2 score at lag time $0.5$~ns and find that backbone torsions contain more kinetic variance than the backbone's heavy atom positions or distances between them (Fig.~\ref{fig:io-to-ck}b).
We note that recently, deep learning approaches for feature selection have been developed that may eventually replace the feature selection step~\cite{vampnet,tae,hernandez-vde}.

Subsequently, we perform TICA~\cite{tica,kinetic-maps} in order to reduce the dimension from the feature space to a low-dimensional space that can be discretized with higher resolution and better statistical efficiency. TICA is a special case of the variational principle~\cite{noe-vac,nueske-vamk} and is designed to find a projection preserving the long-timescale dynamics. Here, TICA of the $\cos/\sin$-transformed backbone torsions at lag time $0.5$ ns yields a four dimensional subspace using a $95\%$ kinetic variance cutoff.
The sample density projected onto the first two independent components (Fig.~\ref{fig:io-to-ck}c) exhibits several maxima, and a time series visualization of said components from the first trajectory (Fig.~\ref{fig:io-to-ck}d) shows discrete jumps between these density maxima.
We thus assume that our transformed feature selection describes one or more metastable processes.

We then proceed with a $k$-means-based discretization of the four dimensional TICA subspace using $k=75$ cluster centers and perform the first validation test, i.e., the estimation of implied timescales (ITS) $t_i$, $i=0, 1,\dots$~\cite{swope-its}
(we note that the number of cluster centers can additionally be chosen using the VAMP score).
The ITS $t_i$ approximates the decorrelation time of the $i^\textrm{th}$ process and is computed from the eigenvalues $\lambda_i$ of the MSM transition matrix via
\begin{equation}
\label{eq:its}
t_i = \frac{-\tau}{\ln\left|\lambda_i(\tau)\right|}
\end{equation}
with $\tau$ being the lag time.
In this validation test, we look for ITS convergence and choose $\tau$ accordingly, i.e., within a range where the ITS are approximately invariant; the uncertainty of the implied timescales is quantified based upon Markov models sampled according to a Bayesian scheme~\cite{ben-rev-msm,noe-tmat-sampling}.
In our example, we find the three slowest ITS to converge quickly and to be constant within a $95\%$ confidence interval for lag times above $0.5$ ns (Fig.~\ref{fig:io-to-ck}e).
We can further see a converged fourth process (cyan line); however, this process decays for lag times above $2.5$ ns and thus might not be well resolved in the final model.

Finally, we estimate a (Bayesian) MSM at a the lag time $\tau=0.5$ ns and validate the model using a Chapman-Kolmogorov (CK) test, which compares the right and the left side of the Chapman-Kolmogorov equation
\begin{equation}
\label{eq:ck}
T(k \tau) = T^k(\tau)
\end{equation}
with $T$ being the MSM transition matrix.
In practice, we utilize a coarse-grained approximation of $T$ for this test, obtained from PCCA++~\cite{pcca++}.
To perform this test, PyEMMA automatically estimates additional MSM transition matrices at lag times $k \tau$ and compares these estimates to the $k^\textrm{th}$ power of the original transition matrix $T(\tau)$.
As we have identified three slow relaxation timescales in the ITS test,
we perform the CK test on the four metastable states whose kinetics involves these three relaxation timescales.
The CK test shows indeed that our model estimated at lag time tau predicts the kinetic behavior on longer timescales accurately (Fig.~\ref{fig:io-to-ck}f).

\subsection{Analyzing the MSM}

\begin{figure}
\includegraphics{figure_3}
\caption{Exemplary analysis of the conformational dynamics of a pentapeptide backbone:
(a) The reweighted free energy surface projected onto the first two independent components exhibits five minima which
(b) PCCA++ identifies as four metastable states.
(c) The second right eigenvector shows that the slowest process shifts probability between the least probable state ($\mathcal{S}_1$) and the two most probable states ($\mathcal{S}_3$, $\mathcal{S}_4$), whereas
(d) the commitor $\mathcal{S}_2\to\mathcal{S}_3$ indicates that the most probable state $\mathcal{S}_4$ acts as a transition state for transitions between states $\mathcal{S}_2$ and $\mathcal{S}_3$.
(e) The Trp-1 SASA shows high variance for all macrostates except $\mathcal{S}_1$ in which the Trp-1 shows the lowest exposure to the solvent.
(f) The Trp-1 SASA auto-correlation function yields a weak signal (top) which, however, can be enhanced if the system is prepared in the non-equilibrium condition $\mathcal{S}_1$ (bottom).
(g) Super-imposed cartoon representations of $200$ molecular conformations sampled from each of the four macrostates show that the backbone appears to obstruct the Trp-1 exposure to the solvent in $\mathcal{S}_{(1,2)}$ but less so in $\mathcal{S}_{(3,4)}$.}
\label{fig:msm-analysis}
\end{figure}

We can now directly extract several thermodynamic and kinetic properties from the estimated and validated model.
An example of the former is the free energy surface in the projection onto the first two TICA components (Fig.~\ref{fig:msm-analysis}a) reweighted based on the MSM stationary distribution.

A fuzzy spectral clustering using the PCCA++ algorithm~\cite{pcca++} allows us to coarse grain the $k$-means microstates into four metastable macrostates (Fig.~\ref{fig:msm-analysis}b) $\mathcal{S}_i$, $i=1,\dots,4$, for which we then approximate the absolute probabilities and relative free energies (defined up to an additive constant)
\[ \begin{array}{ccc}
\textrm{macrostate } \mathcal{S}_i & \pi_{\mathcal{S}_i} & G_{\mathcal{S}_i} / \textrm{k}_\textrm{B} T \\
\hline
\mathcal{S}_1 & 0.004 & 5.565 \\
\mathcal{S}_2 & 0.013 & 4.314 \\
\mathcal{S}_3 & 0.021 & 3.864 \\
\mathcal{S}_4 & 0.962 & 0.039
\end{array}\]
using the relation
\begin{equation}
\label{eq:fe}
G_{\mathcal{S}_i} = - \textrm{k}_\textrm{B} T \ln \sum\limits_{j\in \mathcal{S}_i} \pi_j,
\end{equation}
where $\pi_j$ denotes the MSM stationary weight of the $j^\textrm{th}$ microstate.

In order to interpret the slowest relaxation timescales, we refer to the (right) eigenvectors of the MSM as they contain information about what configurational changes are happening on which timescales.
The first right eigenvector corresponds to the stationary process and its eigenvalue is the Perron eigenvalue $1$.
The second right eigenvector, however, corresponds to the slowest process.
Its minimal and maximal components indicate the microstates between which the process shifts probability.
The relaxation timescale of this exchange process is exactly the corresponding implied timescale.
In the projection onto the first two TICA components, we identify the slowest MSM process as a probability shift between macrostate $\mathcal{S}_1$ and the rest of the system, with macrostates $\mathcal{S}_3$ and $\mathcal{S}_4$ in particular (Fig.~\ref{fig:msm-analysis}c).

The mean first passage times (MFPTs) out and into the macrostate $\mathcal{S}_1$ compute to
\[ \begin{array}{crcr}
\textrm{direction} & \textrm{mean / ns} && \textrm{std / ns} \\
\hline
\mathcal{S}_1 \to \mathcal{S}_{(2,3,4)} & 8.9 & \pm & 1.8 \\
\mathcal{S}_{(2,3,4)} \to \mathcal{S}_1 & 2487.5 & \pm &  504.7
\end{array}\]
using the Bayesian MSM.

TPT~\cite{weinan-tpt,metzner-msm-tpt} is method to analyze the statistics of transition pathways. The TPT version of~\cite{noe-folding-pathways} can be conveniently applied to the estimated MSM. Here, we compute the TPT flux between macrostates $\mathcal{S}_2$ and $\mathcal{S}_3$ (Fig.~\ref{fig:msm-analysis}d).
The committor projection onto the first two TICA components shows that it is constant within the metastable states defined above.
Transition regions (macrostate $\mathcal{S}_4$) can be identified by committor values $\approx \frac{1}{2}$.

To illustrate how to connect an MSM analysis with experiments, we compute the Trp-1 solvent accessible surface area (SASA) using the mdtraj~\cite{mdtraj} package.
The SASA is a reasonable approximation for the instantaneous tryptophan flourescence signal.
The auto correlation function (ACF) of this signal can be measured in spectroscopy experiments.

In the projection onto the first two TICA components, we observe a high variance of the SASA in all macrostates except $\mathcal{S}_1$ (Fig.~\ref{fig:msm-analysis}e).
This indicates that Trp-1 has only a low solvent exposure in this macrostate.
The ACF itself yields a weak signal (Fig.~\ref{fig:msm-analysis}f, top panel).
If, however, we prepare the system in macrostate $\mathcal{S}_1$ (non-equilibrium condition), the relaxation towards equilibrium yields a much stronger signal (Fig.~\ref{fig:msm-analysis}f, bottom panel).

Finally, we sample $200$ super-imposed molecular conformations from each of the $4$ macrostates (Fig.~\ref{fig:msm-analysis}g) and render them using VMD~\cite{vmd}.
We observe that, in macrostates $\mathcal{S}_1$ and $\mathcal{S}_2$, the backbone appears to often be in close proximity to the Trp-1 (shown in red), thus effectively limiting Trp-1 exposure to the solvent.
In the (more probable) macrostates $\mathcal{S}_3$ and $\mathcal{S}_4$, the Trp-1 appears to be much more exposed.

\subsection{Summary}

In this section, we have summarized how to conduct an MSM-based analysis of biomolecular dynamcis data using PyEMMA. For the full analysis, please refer to the first notebook (00). All notebooks as well as detailed installation instructions are available on \githubrepository{}.

\section{Author Contributions}
%%%%%%%%%%%%%%%%
% This section mustt describe the actual contributions of
% author. Since this is an electronic-only journal, there is
% no length limit when you describe the authors' contributions,
% so we recommend describing what they actually did rather than
% simply categorizing them in a small number of
% predefined roles as might be done in other journals.
%
% See the policies ``Policies on Authorship'' section of https://livecoms.github.io
% for more information on deciding on authorship and author order.
%%%%%%%%%%%%%%%%
CW, MKS, TH, SO, and FN designed research.
CW, MKS, TH, SO, and BHE developed and tested notebooks.
MKS developed the software infrastructure, test, and install environment.
CW, MKS, TH, BEH, and FN wrote the manuscript.

For a more detailed description of author contributions, see the GitHub issue tracking and changelog at\\\githubrepository{}.

\section{Other Contributions}
%%%%%%%%%%%%%%%
% You should include all people who have filed issues that were
% accepted into the paper, or that upon discussion altered what was in the paper.
% Multiple significant contributions might mean that the contributor
% should be moved to authorship at the discretion of the a
%
% See the policies ``Policies on Authorship'' section of https://livecoms.github.io for
% more information on deciding on authorship and author order.
%%%%%%%%%%%%%%%
We are grateful to Nuria Plattner for providing the pentapeptide simulation data and Camilla Ventura Santos as well as the entire computational molecular biology group for valuable discussion and feedback.

For a more detailed description of contributions from the community and others, see the GitHub issue tracking and changelog at \githubrepository{}.

\section{Potentially Conflicting Interests}
%%%%%%%
%Declare any potentially competing interests, financial or otherwise
%%%%%%%
The authors declare no conflicting interests.

\section{Funding Information}
%%%%%%%
% Authors should acknowledge funding sources here. Reference specific grants.
%%%%%%%
TH acknowledges financial support by SFB/TRR 186. SO acknowledges a postdoctral fellowship from the Alexander von Humboldt Stiftung.

\bibliography{literature}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% APPENDICES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\appendix

\end{document}